---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r, echo = FALSE, results='hide'}
library(data.table)
library(ggplot2)
library(lattice)
library(dplyr)
library(timeDate)
```
  
## Loading and preprocessing the data
  
```{r}
setwd("~/ScriptsR/RepData_PeerAssessment1")
Activity <- read.csv("activity.csv", sep = ",",
                       header = TRUE, na.strings = "NA")
Activity$date <- as.Date(Activity$date, "%Y-%m-%d")
names(Activity)
class(Activity)
dim(Activity)
```


## What is mean total number of steps taken per day?
  
Total steps per day:
```{r}
by_day_Activity <- group_by(Activity, date)
Activity_by_day <- as.data.frame(summarise(by_day_Activity, sum.Steps = sum(steps, na.rm = TRUE)))
g1<-ggplot(Activity_by_day,aes(date, sum.Steps))
g1 + geom_bar(stat = "identity") + labs(y = "Steps") + labs(x = "Date") + 
    labs(title="Total Steps per Day")+theme(plot.title = element_text(hjust = 0.5))
```
  
Mean and median steps per day:
```{r}
Activity_by_day <- group_by(Activity, date) %>% summarise(Mean.Steps = mean(steps, na.rm = TRUE), Median.Steps = median(steps, na.rm = TRUE))
Activity_by_day <- as.data.frame(Activity_by_day)
Activity_by_day

```
  
## What is the average daily activity pattern?
  
Mean steps per day:
```{r}
by_day_Activity <- group_by(Activity, date)
Activity_by_day <- as.data.frame(summarise(by_day_Activity, Mean.Steps = mean(steps, na.rm = TRUE)))
g2<-ggplot(Activity_by_day,aes(date,Mean.Steps))
g2 + geom_line() + labs(y = "Mean of Steps") + labs(x = "Date") + 
    labs(title="Mean Steps per Day")+theme(plot.title = element_text(hjust = 0.5))
```
  
5-minute interval with maximum number of steps:  
```{r}
ActivityInt<-Activity
ActivityInt$interval <- as.factor(ActivityInt$interval)
Activity_by_interval <- group_by(ActivityInt, interval) %>% summarise(Mean.Steps = mean(steps, na.rm = TRUE))
Activity_by_interval <- as.data.frame(Activity_by_interval)
Activity_by_interval <- filter(Activity_by_interval, Mean.Steps == max(Mean.Steps))
Activity_by_interval
``` 

## Imputing missing values

Replacing NA by the mean of the 5-minute interval:  

```{r}
StepsAct<-Activity$steps
summary(StepsAct)
mean(is.na(StepsAct))
by_interval <- group_by(Activity, interval)
Activity_by_interval <- as.data.frame(summarise(by_interval, Mean.Steps = mean(steps, na.rm = TRUE)))
ActivityInt <- merge(Activity, Activity_by_interval, by= "interval")
ActivityInt$steps <- ifelse (is.na(ActivityInt$steps), ActivityInt$Mean.Steps, ActivityInt$steps)
ActivityInt <- arrange(ActivityInt, date, interval)
StepsAct<-ActivityInt$steps
summary(StepsAct)
``` 
  
Mean and meadia after replacing NA:     
```{r}
Activity_by_day <- group_by(ActivityInt, date) %>% summarise(Mean.Steps = mean(steps, na.rm = TRUE), Median.Steps = median(steps, na.rm = TRUE))
Activity_by_day <- as.data.frame(Activity_by_day)
Activity_by_day
```

Histograma after replacing NA:  
```{r}
by_day_Activity <- group_by(ActivityInt, date)
Activity_by_day <- as.data.frame(summarise(by_day_Activity, sum.Steps = sum(steps, na.rm = TRUE)))
g1<-ggplot(Activity_by_day,aes(date, sum.Steps))
g1 + geom_bar(stat = "identity") + labs(y = "Steps") + labs(x = "Date") + 
    labs(title="Total Steps per Day")+theme(plot.title = element_text(hjust = 0.5))
```

## Are there differences in activity patterns between weekdays and weekends?
  
Plot with difference between weekdays and weekends:  
```{r}
ActivityWeek<-Activity
ActivityWeek$week<-ifelse(isWeekday(ActivityWeek$date),"WeekDay","Weekend")
ActivityWeek<-transform(ActivityWeek, week=factor(week))
by_weekday_Activity <- group_by(ActivityWeek,week,interval)
Activity_by_weekday <- as.data.frame(summarise(by_weekday_Activity, Mean.Steps = mean(steps, na.rm = TRUE)))
g<-ggplot(Activity_by_weekday, aes(interval,Mean.Steps))
g+facet_grid(week~.)+geom_line()
``` 
  
